$\pi$ with leftovers: \\ a mechanisation in Agda

### 2021-01-26 paper/appendix.tex
Apply remaining fixes

\begin{proof}[Proof]
  All the above results are proved by induction on \textsc{Process}, \textsc{Var}, and \textsc{Raw} (\autoref{def:syntax}).
  Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
\end{proof}

@@

\begin{proof}[Proof]
  All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
\end{proof}

@@

\begin{proof}[Proof]
  The proof is by induction on \textsc{Equals} $\eq{}$.
  Here we only consider those cases that are not purely inductive: the base cases for $\constr{struct}$ and their symmetric variants.
  Full proof in \cite{Zalakain2020Agda}.
  We proceed by induction on \textsc{StructCong} $\eqeq$:
  \begin{itemize}
    \item
      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
    \item
      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
    \item
      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
    \item
      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$.
    \item
      Case $\constr{scope-ext}$ for $\new (\comp{P}{Q})$: we need to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
    \item
      Case $\constr{scope-comm}$:
      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged. \qed
  \end{itemize}
\end{proof}

@@

\begin{nitheorem}[Generalised substitution]
  \label{thm:substitution-generalization}
  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  The substituted variable $i$ is capable of $m$ in $\Gamma_i$, and capable of $n$ in $\Psi_i$.
  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
  In other words, there must exist some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
  Then substituting $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
  Refer to \autoref{fig:subst} for a diagramatic representation.
\end{nitheorem}

\begin{proof}[Proof]
  By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  \begin{itemize}
    \item
      For constructor $\PO$ we get $\Gamma_i \equiv \Psi_i$.
      From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
      Therefore $\Gamma_j \equiv \Psi_j$ and $\constr{end}$ can be applied.

@@

      \begin{minipage}[b]{0.5\linewidth}
        \centering
        \begin{tikzpicture}
          \node (gamma-i) at (0,4)   {$\Gamma_i$};
          \node (gamma-m) at (0,2)   {$\Gamma$};
          \node (gamma-j) at (0,0)   {$\Gamma_j$};
          \node (xi-i)    at (3,3.5) {$\Xi_i$};
          \node (theta)   at (3,2.3) {$\Theta$};
          \node (delta-m) at (3,1.7) {};
          \node (xi-j)    at (3,0.5) {$\Xi_j$};
          \node (psi-i)   at (6,3)   {$\Psi_i$};
          \node (psi-m)   at (6,2)   {$\Psi$};
          \node (psi-j)   at (6,1)   {$\Psi_j$};

@@

        Diagrammatic representation of the $\comp{}{}$ case for substitution.
        Continuous lines represent known facts, dotted lines proof obligations.
      \end{minipage}
      \begin{minipage}[b]{0.5\linewidth}
        \centering
        \begin{tikzpicture}
          \node (gamma-i) at (0,2)   {};
          \node (gamma-m) at (0,1)   {};
          \node (gamma-j) at (0,0)   {};
          \node (xi-i)    at (2,3)   {};
          \node (xi-m)    at (2,2) {};
          \node (xi-j)    at (2,1)   {};
          \node (psi-m)   at (4,2) {};

          \draw[->] (gamma-i) -- (xi-j);
          \draw[->] (gamma-m) -- (xi-j);
          \draw[->,densely dotted] (gamma-j) -- (xi-j);
          \draw[->] (gamma-i) -- (gamma-m);
          \draw[->] (gamma-j) -- (gamma-m);
          \draw[->] (xi-i)    -- (xi-m);
          \draw[->] (xi-j)    -- (xi-m);
          \draw[->] (xi-i)    -- (psi-m);
          \draw[->] (xi-m)    -- (psi-m);
          \draw[->,densely dotted] (xi-j)    -- (psi-m);
        \end{tikzpicture}

        Alternative approach without arrows $\ni_i n$ and $\ni_j n$.
        This approach makes composing typing derivations before and after subsitution more difficult.
      \end{minipage}

@@

\begin{nitheorem}[Substitution]
  \label{thm:substitution}
  Let process $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
\end{nitheorem}
\begin{proof}[Proof]
  For $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$, we use framing to derive them.
  Then, we use these to apply \autoref{thm:substitution-generalization}.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{note}
  We use $\containsusage{\Gamma}{i}{x}{\Delta}$ to stand for $\contains{\gamma}{\Gamma}{i}{t}{x}{\Delta}$ for some $\gamma$ and $t$.
% \end{note}

@@

\begin{proof}[Proof]
  By induction on $P \reduce{c} Q$. For the full details refer to our mechanisation in Agda.  
  \begin{itemize}
    \item
    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).

>>>>

\begin{proof}
  All the above results are proved by induction on \textsc{Process}, \textsc{Var} (\autoref{def:syntax}) and \textsc{Raw}.
  Complete details can be found in our mechanisation in Agda in \cite{Zalakain2020Agda}.
\end{proof}

@@

\begin{proof}
  All the above theorems are proved by induction on \textsc{Types} and \textsc{VarRef}.
  For details, refer to our mechanisation in Agda \cite{Zalakain2020Agda}.
\end{proof}

@@

\begin{proof}
  The proof is by induction on \textsc{Equals} $\eq{}$.
  Here we only consider those cases that are not purely inductive: the base cases for $\constr{struct}$ and their symmetric variants.
  Full proof in \cite{Zalakain2020Agda}.
  We proceed by induction on \textsc{StructCong} $\eqeq$:
  \begin{itemize}
    \item
      Case $\constr{comp-assoc}$: trivial, as leftover typing is naturally associative.
    \item
      Case $\constr{comp-sym}$ for $\comp{P}{Q}$: we use framing (\autoref{thm:framing}) to shift the output context of $P$ to the one of $Q$; and the input context of $Q$ to the one of $P$.
    \item
      Case $\constr{comp-end}$: trivial, as the typing rule for $\PO$ has the same input and output contexts.
    \item
      Case $\constr{scope-end}$: we show that the usage annotation of the newly created channel must be $\lz$, making the proof trivial.
      In the opposite direction, we instantiate the newly created channel to a type $\unit$ and a usage annotation $\lz$.
    \item
      Case $\constr{scope-ext}$ for $\new (\comp{P}{Q})$: we need to show that $P$ preserves the usage annotations of the unused variable (\autoref{lm:types-unused}) and then use strengthening (\autoref{thm:strengthening}).
      In the reverse direction, we use weakening (\autoref{thm:weakening}) on $P$ and show that lowering and then lifting $P$ results in $P$ (\autoref{lm:lower-lift}).
    \item
      Case $\constr{scope-comm}$:
      we use exchange (\autoref{thm:exchange}), and for the reverse direction exchange and \autoref{lm:exchange-exchange} to show that exchanging two elements in $P$ twice leaves $P$ unchanged. \qed
  \end{itemize}
\end{proof}

@@

\begin{nitheorem}[Generalised substitution]
  \label{thm:substitution-generalization}
  Let process $P$ be well-typed in $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  The substituted variable at position $i$ can be split into $m$ in $\Gamma_i$, and into $n$ in $\Psi_i$.
  Substitution will take these usages $m$ and $n$ away from $i$ and transfer them to the variable $j$ we are substituting for.
  In other words, let there be some $\Gamma$, $\Psi$, $\Gamma_j$ and $\Psi_j$ such that:
  \begin{multicols}{2}
  \begin{itemize}
    \item $\contains{\gamma}{\Gamma_i}{i}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Gamma_j}{j}{t}{m}{\Gamma}$
    \item $\contains{\gamma}{\Psi_i}{i}{t}{n}{\Psi}$
    \item $\contains{\gamma}{\Psi_j}{j}{t}{n}{\Psi}$
  \end{itemize}
  \end{multicols}
  Let $\Gamma$ and $\Psi$ be related such that $\opctx{\Gamma}{\Delta}{\Psi}$ for some $\Delta$.
  Let $\Delta$ have a usage annotation $\lz$ at position $i$, so that all consumption from $m$ to $n$ must happen in $P$.
  Then substituting $i$ to $j$ in $P$ will be well-typed in $\types{\gamma}{\Gamma_j}{\subst{P}{j}{i}}{\Psi_j}$.
\end{nitheorem}

\begin{proof}
  By induction on the derivation $\types{\gamma}{\Gamma_i}{P}{\Psi_i}$.
  \begin{itemize}
    \item
      For constructor $\PO$ we get $\Gamma_i \equiv \Psi_i$.
      From $\Delta_i \equiv \lz$ follows that $m \equiv n$.
      Therefore $\Gamma_j \equiv \Psi_j$ and $\constr{end}$ can be applied.

@@

      \centering
      \begin{tikzpicture}
        \node (gamma-i) at (0,4)   {$\Gamma_i$};
        \node (gamma-m) at (0,2)   {$\Gamma$};
        \node (gamma-j) at (0,0)   {$\Gamma_j$};
        \node (xi-i)    at (3,3.5) {$\Xi_i$};
        \node (theta)   at (3,2.3) {$\Theta$};
        \node (delta-m) at (3,1.7) {};
        \node (xi-j)    at (3,0.5) {$\Xi_j$};
        \node (psi-i)   at (6,3)   {$\Psi_i$};
        \node (psi-m)   at (6,2)   {$\Psi$};
        \node (psi-j)   at (6,1)   {$\Psi_j$};

@@

      Diagrammatic representation of the $\comp{}{}$ case for substitution.
      Continuous lines represent known facts, dotted lines proof obligations.

@@

\begin{nitheorem}[Substitution]
  \label{thm:substitution}
  Let process $P$ be well typed in $\types{\gamma \comma t}{\Gamma \comma m}{P}{\Psi \comma \lz}$.
  Let $\contains{\gamma}{\Psi}{j}{t}{m}{\Xi}$.
  Then, we can substitute the variable references to $\constr{0}$ in $P$ with $\suc j$ so that the result is well typed in $\types{\gamma \comma t}{\Gamma \comma m}{\subst{P}{\suc j}{\constr{0}}}{\Xi \comma m}$.
\end{nitheorem}
\begin{proof}
  For $\contains{\gamma}{\Gamma}{j}{t}{m}{\Theta}$ and $\types{\gamma \comma t}{\Theta \comma m}{P}{\Xi \comma \lz}$ for some $\Theta$, we use framing to derive them.
  Then, we use these to apply \autoref{thm:substitution-generalization}.
\end{proof}

@@

\begin{proof}
  By induction on $P \reduce{c} Q$. For the full details refer to our mechanisation in Agda.  
  \begin{itemize}
    \item
    Case $\constr{comm}$: we apply framing (\autoref{thm:framing}) (to rearrange the assumptions), substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).
